The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories

### 2017-09-19 JStatSoft_forestinventory_master-concordance.tex
Infinite pop approach

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 107 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 115:%
1 97 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 213:%
123 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 223:%
1 131 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 355:%
134 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 367:%
1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 981:%
147}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 982:%
1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1501:%
149 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1513:%
1 42 1 1 2 11 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1578:%
162 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1588:%
1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1885:%
173 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1895:%
1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2126:%
184 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2135:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2140:%
194 11 1}

>>>>

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 107 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 115:%
1 97 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 213:%
123 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 223:%
1 132 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 356:%
134 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 368:%
1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1004:%
147}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 1005:%
1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1524:%
149 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1536:%
1 42 1 1 2 11 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1601:%
162 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1611:%
1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1908:%
173 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1918:%
1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2149:%
184 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2158:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2163:%
194 11 1}

### 2017-09-19 rnw/Main structure and modules of the Package.Rnw
Infinite pop approach

Most forest inventories typically follow a procedure that involves randomly selecting points in the forest area defined by GPS coordinates to which inventorists are sent. At these points, the inventorist demarcates a sample plot, possibly with the help of an inclusion circle that indicates which trees above a certain threshold are to be selected for measurement. These measurements are then converted into local plot densities (e.g. the timber density in $\frac{m^3}{ha}$) before the desired estimator is applied.  The \pkg{forestinventory} package bridges this common inventory paradigm to the mathematics of the estimation process using what is called the \textit{infinite population approach}.  In this approach the sampling units are \textit{points} each of which has a local density that exists but might remain unobserved if that point was not selected in our inventory's sample. For the sake of argument, imagine that the local density for every point in the forested area was somehow known and we could visualize the spatial distribution of the local density function of a target variable (e.g. the timber volume) for the entire forest area (Figure \ref{fig:inf_pop_apr}). The \textit{population total} of the target variable (e.g. the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which defines the volume under the green surface in (Figure \ref{fig:inf_pop_apr}). Clearly, evaluating the local densities for every one of the infinite point so that we can add them up is impossible. Thus, in the infinite population approach we are essentially approximating the integral of this unobserved density surface numerically by sampling directly from the surface with our terrestrial inventory sample. This is why the infinite population approach is also called the \textit{Monte-Carlo approach}.

All estimators included in the \pkg{forestinventory} package basically rest upon these theoretical thoughts. The motivation for this decision was that no bothersome spatial correlations need to be considered as would be the case with geostatisical methods since we assume that the inventory points are selected uniformly and independently, i.e. the infinite analogy for simple random sampling. Furthermore, points can perfectly fill up any forest area with any amorphous shape, which is not possible if one were to use the \textit{finite populatoin approach} where the sampling plot defined around the point is considered the sampling unit.  However, this has the disadvatage that circles don't tesselate perfectly and polygons will inevitably distort and pixelate the forest boundary.  In practice, the population of sampling units in the finite approach will almost never perfectly cover the forests.  We also want to emphasize that the consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.

@@

The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly referred to as \textit{phases} (Figure \ref{fig:concmphase_and_sae_a}). The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$. The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e. the local density $Y(x)$ (e.g. the timber volume density per hectare). The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS) technique. The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $x \in s_1$ and to correct any possible model biases, i.e. $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed. In the design-based context, the two-phase estimator is unbiased regardless of the model used to produce the predictions.  This property comes from the assumption that each phase's sample is selected via simple random sampling (see section \ref{sec:db_vs_md}).
%In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.

>>>>

Most forest inventories typically follow a procedure that involves randomly selecting points in the forest area defined by GPS coordinates to which inventorists are sent. At these points, the inventorist demarcates a sample plot, possibly with the help of an inclusion circle that indicates which trees above a certain threshold are to be selected for measurement. These measurements are then converted into local plot densities (e.g. the timber density in $\frac{m^3}{ha}$) before the desired estimator is applied.  The \pkg{forestinventory} package bridges this common inventory paradigm to the mathematics of the estimation process using what is called the \textit{infinite population approach}.  In this approach the sampling units are \textit{points} each of which has a local density that exists but might remain unobserved if that point was not selected in our inventory's sample. For the sake of argument, imagine that the local density for every point in the forested area was somehow known and we could visualize the spatial distribution of the local density function of a target variable (e.g. the timber volume) for the entire forest area (Figure \ref{fig:inf_pop_apr}). The \textit{population total} of the target variable (e.g. the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which defines the volume under the green surface in Figure \ref{fig:inf_pop_apr}. Practically speaking, attempting to evaluate the local densities for every one of the infinite points so that we can add them up is impossible. Thus, in the infinite population approach we are essentially approximating the integral of this unobserved density surface numerically by sampling directly from the surface with our terrestrial inventory sample. This is why the infinite population approach is also called the \textit{Monte-Carlo approach}.

All estimators included in the \pkg{forestinventory} package basically rest upon this theoretical approach. The motivation for this decision was that no bothersome spatial correlations need to be considered, as would be the case with geostatisical methods, since we assume that the inventory points are selected uniformly and independently, i.e. the infinite analogy for simple random sampling. Furthermore, points can perfectly fill up any forest area with any amorphous shape, which is not possible if one were to use the \textit{finite population approach} where the sampling plot defined around the point is considered the sampling unit. The finite population approach has the disadvantage that circles don't tesselate perfectly and polygons will inevitably distort and pixelate the forest boundary.  In practice, the population of sampling units in the finite approach will almost never perfectly cover the forests. The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.

@@

The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly referred to as \textit{phases} (Figure \ref{fig:concmphase_and_sae_a}). The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$. The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e. the local density $Y(x)$ (e.g. the timber volume per hectare). The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS) technique. The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $x \in s_1$ and to correct any possible model biases, i.e. $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed. In the design-based context, the two-phase estimator is unbiased regardless of the model used to produce the predictions.  This property comes from the assumption that each phase's sample is selected via simple random sampling (see section \ref{sec:db_vs_md}).
%In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.
